sofa
=======

```{r, echo=FALSE}
knitr::opts_chunk$set(
  collapse=TRUE,
  comment="#>",
  warning=FALSE, 
  message=FALSE
)
```

<pre>
  _ _ _ _ _ _ _ _ _ _ _ 
 /|                   |\
/ |_ _ _ _ _ _ _ _ _ _| \
\ /                    \/
 \ ___________________ /
</pre>

#### *An easy interface to CouchDB from R*

Note: Check out [*R4couchdb*](https://github.com/wactbprot/R4CouchDB), another R package to interact with CouchDB. 

## Quickstart

### Install CouchDB

Instructions [here](http://wiki.apache.org/couchdb/Installation)

### Connect to CouchDB

In your terminal 

```sh
couchdb
```

You can interact with your CouchDB databases as well in your browser. Navigate to [http://localhost:5984/_utils](http://localhost:5984/_utils)

### Install sofa

From GitHub

```{r eval=FALSE}
install.packages("devtools")
devtools::install_github("sckott/sofa")
```

```{r}
library('sofa')
```

### Cushions

Cushions? What? Since it's couch we gotta use `cushions` somehow. `cushions` are basically just a simple named list holding details of connections for different couches you work with. See `?cushions` or `?authentication` for help. 

As an example, here's how I set up details for connecting to my Cloudant couch:

```{r}
cushion(name = 'cloudant', user = '<user name>', pwd = '<password>', type = "cloudant")
```

Break down of parameters: 

* `name`: Name of the cushion. This is how you'll refer to each connection. `cushion` is the first parameter of each function. 
* `user`: User name for the service.
* `pwd`: Password for the service, if any.
* `type`: Type of cushion. This is important. Only `localhost`, `cloudant`, and `iriscouch` are supported right now. Internally in `sofa` functions this variable determines how urls are constructed for http requests. 
* `port`: The port to connect to

Of course by default there is a built in `cushion` for localhost so you don't have to do that, unless you want to change those details, e.g., the port number.

Right now cushions aren't preserved across R sessions, but working on that.

### Ping the server

```{r}
ping()
```

Nice, it's working.

### Create a new database, and list available databases

```{r echo=FALSE}
deletedb(dbname='sofadb')
```

```{r}
createdb(dbname='sofadb')
```

see if its there now

```{r}
listdbs()
```

### Create documents

#### Write a document WITH a name (uses PUT)

```{r}
doc1 <- '{"name":"sofa","beer":"IPA"}'
doc_create(dbname="sofadb", doc=doc1, docid="a_beer")
```

#### Write a json document WITHOUT a name (uses POST)

```{r}
doc2 <- '{"name":"sofa","icecream":"rocky road"}'
doc_create(dbname="sofadb", doc=doc2)
```

#### XML? 

Write an xml document WITH a name (uses PUT). The xml is written as xml in couchdb, just wrapped in json, when you get it out it will be as xml.

write the xml

```{r}
doc3 <- "<top><a/><b/><c><d/><e>bob</e></c></top>"
doc_create(dbname="sofadb", doc=doc3, docid="somexml")
```

get the doc back out

```{r}
doc_get(dbname="sofadb", docid="somexml")
```

get just the xml out

```{r}
doc_get(dbname="sofadb", docid="somexml")[["xml"]]
```

### Views

First, create a database

```{r}
createdb(dbname='alm_couchdb')
```

Write a view - here letting key be the default of null

```{r}
view_put(dbname='alm_couchdb', design_name='almview2', value="doc.baseurl")
```

get info on your new view

```{r}
view_get(dbname='alm_couchdb', design_name='almview2')
```

get data using a view

```{r}
view_search(dbname='alm_couchdb', design_name='almview2')
```

delete the view

```{r}
view_del(dbname='alm_couchdb', design_name='almview2')
```


### Full text search? por sepuesto

I'm working on an R client for Elaticsearch called `elastic` - find it at [https://github.com/ropensci/elastic](https://github.com/ropensci/elastic)

Thinking about where to include functions to allow `elastic` and `sofa` to work together...if you have any thoughts hit up the issues. 
