<!--
%\VignetteEngine{knitr}
%\VignetteIndexEntry{An R Markdown Vignette made with knitr}
-->
  
sofa - easy interface to CouchDB from R
======
  
### What is sofa?
  
XXXXXXX

It is getting easier to get data directly into R from the web. Often R packages that retrieve data from the web return useful R data structures to users like a data.frame. This is a good thing of course to make things user friendly. 

However, what if you want to drill down into the data that's returned from a query to a database in R?  What if you want to get that nice data.frame in R, but you think you may want to look at the raw data later? The raw data from web queries are often JSON or XML data. This type of data, especially JSON, can be easily stored in schemaless so-called NoSQL databases, and queried later. 

There are plenty of databases you can interact with from R, so why CouchDB? For one, it makes a lot of sense to write to a NoSQL database when you have JSON, XML, HTML, etc., which aren't a good fit for databases like MySQL, SQLite, PostgreSQL, etc. ([though postgres allows you to write JSON][postgres]). It didn't have to be CouchDB, but at least to me it seems relatively easy to install, you can interact with it via an HTTP API (if you're into that, which I am), and it has a nice web interface (navigate to [http://localhost:5984/_utils/](http://localhost:5984/_utils/) after starting `couchdb`).

### A brief aside: JSON and XML

What are JSON and XML? This is what JSON looks like (ps if you ever wonder if your JSON is correct, go [here](http://jsonlint.com/)):

```bash
{
  "name": "joe",
  "hobby": "codemonkey",
  "lives": [
      {
          "city": "San Jose",
          "state": "CA"
      }
  ]
}
```

This is what XML looks like:

```bash
<?xml version="1.0" encoding="UTF-8" ?>
  <name>joe</name>
  <hobby>codemonkey</hobby>
	<lives>
		<city>San Jose</city>
		<state>CA</state>
	</lives>
```

### CouchDB help

+ [Apache CouchDB](http://couchdb.apache.org/).
+ [CouchDB guide - online book and hard copy](http://guide.couchdb.org/).
+ [Couchapp - sort of like rails, for templating apps based on CouchDB](http://couchapp.org/page/index).

### Quick start

#### Install sofa

```{r install, comment=NA, warning=FALSE}
# install.packages("devtools"); library(devtools); install_github("rbison", "ropensci")
library(sofa)
```

#### Ping your local server just to check your head
```{r ping, comment=NA, warning=FALSE}
sofa_ping()
```

### More in depth stuff

#### Working with documents

#### Working with a remote database

#### Attachements

#### The changes feed